I.Sicily. An EpiDoc corpus for ancient Sicily

### 2015-12-17 source/24/Epigraphy and onomastics in Hesperia databank-template.tex
no indent added

The aim of this work is to do a general overview of the computer-based epigraphic corpora, to think about their features
and advantages, taking the experience gained in the last years within the Hesperia project as a starting point.\footnote{ This paper is an output from the FFI2012-25113 project and the Senior Research Team LITTERA
(2014SGR63), on the one hand, and 655938 — ZEPHYROS — H2020-MSCA-IF-2014 project, on the other.}

@@

All above-mentioned principles have been taken into account in Hesperia project, its main goal being to gather all
linguistic evidences from Palaeohispanic languages, that is, pre-Roman languages in the Iberian Peninsula\footnote{
General presentations of the project can be found in \citet{orduna_implementing_????}, \citet{orduna_philology_????} and
\citet{velaza_hesperia:_2014}.}. First of all, we need to recall these are very different materials as for their quantity, quality and
reliability. Ancient Hispania left us epigraphic texts in at least four languages: Tartessian, Iberian, Celtiberian and
Lusitanian -{}-the possibility there are also some texts in Palaeobasque language is still sub iudice\footnote{
Concerning this subject, see \citet{velaza_epigrafiy_2009}.}{}-{}-, mainly written in four epichoric scripts -{}-Tartessian,
southeastern-Iberian, northeastern-Iberian and Greek-Iberian script.

@@

The Hesperia databank has been recently extended with a new resource on indigenous onomastics. More specifically, the
new sections now available are devoted to the anthroponymy and theonymy, leaving for the near future the data referring
to toponymy and ethnonymy, which are not yet available in open access.

>>>>

\noindent The aim of this work is to do a general overview of the computer-based epigraphic corpora, to think about their features
and advantages, taking the experience gained in the last years within the Hesperia project as a starting point.\footnote{ This paper is an output from the FFI2012-25113 project and the Senior Research Team LITTERA
(2014SGR63), on the one hand, and 655938 — ZEPHYROS — H2020-MSCA-IF-2014 project, on the other.}

@@

\noindent All above-mentioned principles have been taken into account in Hesperia project, its main goal being to gather all
linguistic evidences from Palaeohispanic languages, that is, pre-Roman languages in the Iberian Peninsula\footnote{
General presentations of the project can be found in \citet{orduna_implementing_????}, \citet{orduna_philology_????} and
\citet{velaza_hesperia:_2014}.}. First of all, we need to recall these are very different materials as for their quantity, quality and
reliability. Ancient Hispania left us epigraphic texts in at least four languages: Tartessian, Iberian, Celtiberian and
Lusitanian -{}-the possibility there are also some texts in Palaeobasque language is still sub iudice\footnote{
Concerning this subject, see \citet{velaza_epigrafiy_2009}.}{}-{}-, mainly written in four epichoric scripts -{}-Tartessian,
southeastern-Iberian, northeastern-Iberian and Greek-Iberian script.

@@

\noindent The Hesperia databank has been recently extended with a new resource on indigenous onomastics. More specifically, the
new sections now available are devoted to the anthroponymy and theonymy, leaving for the near future the data referring
to toponymy and ethnonymy, which are not yet available in open access.

### 2015-12-17 source/29/EAGLE2016_camera_ready.tex
no indent added

The technological advances in the last decade have equipped the general public with several handheld electronic
commodities that changed significantly daily routine in a personal and professional level and contributed to the user's
quality of life not only in developed countries but also in developing economies in Africa and Asia [Osman 2011].
Handheld devices, such as tablets and smart phones, not only connect the users with tremendous amount of information
through the internet, but also offer interfaces for natural user interaction that enable non-technology-oriented
populations to use computers intuitively.

@@

Understanding the users is one of the integral steps of interaction design, which is an iterative process during which
representative users interact with preliminary designs and provide useful feedback \citep{preece_interaction_2015}. For the
purposes of this study, our team interacted with early adopters of our prototype system, who were epigraphists and
conservation specialists from Cornell University, the University of California, Berkeley, the University of Lyon 2, the
Berlin-Bradenburg Academy of Sciences and Humanities, the U.K. National Archives, and the University of Florida. The
goals of our interaction were twofold: a) to study the various forms of physical interaction that epigraphists have
with an inscription as a real physical object and b) to expose epigraphists to a digital interface that imitates
their interaction routine, using digital replicas of physical objects.

@@

Natural User Interfaces consist of sensors that track the natural behavior of users and provide a natural form of
interactivity with computers and other electronic devices. The common forms of NUI sensors are: pressure sensors for
sensing touch gestures (e.g. touch screens and touch pads), motion sensors for sensing user-initiated changes in the
orientation and acceleration of the device (e.g. accelerometer, gyroscope, and compass), and position sensors for
tracking changes in the relative position of the user with respect to the device, such as body motions (e.g.
Microsoft's Kinect), finger motions (e.g. Occipital's Leap Motion), eye movements, and others.

\ An optimal interaction design solution should be intuitive, minimalistic, and non intrusive \citep{preece_interaction_2015}.
Therefore, in order to design interaction for epigraphy one should choose devices that are easily accessible by
epigraphists and do not interfere with their workspace (e.g. avoid introducing new devices or external sensors). All
forms of interaction described in Sec. 1.2 can be implemented, using motion and pressure sensors, which can be easily
found in tablet computers or smart phones. In both types of handheld devices the virtual object can also be assumed
handheld, without loss of generality, in order to generate a multi-sensory experience for the user (i.e. holding the
device = holding the digital inscription). Hence, NUI design is possible by utilizing accessible devices and
without the use of external sensors as it is described in details in the following sections.

@@

In order to achieve natural interactive relighting of an inscription, the system should imitate the process of
relighting a handheld physical object (such as a paper cast of inscription) by reorienting the object with respect to
the light of the environment. Without loss of generality, we can assume that the default virtual lighting source is
located on the ceiling, right above the device, which is also very intuitive choice as it is the most probable
real-world lighting condition. Under this assumption, a gyroscope, a sensor that tracks the orientation of the device
with respect to the gravitational vector, is enough to track the slope of the device with respect to the virtual light.
The top row of Fig.~\ref{fig:1} shows the approximated real-world orientation of a tablet computer as it was estimated using
the gyroscope of the device. The orientation is updated in real-time as the user moves the device.

@@

Another important form of interaction in the epigraphic routine is the change of the point of view in order to
understand better the structural details of the inscribed letterforms. Assuming that the model of the inscription is
parallel to the screen of the device, the change of point of view involves only change of the perspective projection of
the digital object without any virtual rotation. In such case the rotation of the object is equivalent to the physical
rotation of the device without any virtual rotation of the object.

@@

The forms of interactions presented in 1.3.1 and 1.3.2 involved only the motion sensors of a tabled computer without
utilizing the pressure sensors of the screen of the device. The commonly used touch gestures (such as 2-finger twist to
rotate, 2-finger pinch to zoom, and 2-finger translate to move) can be employed in order to enhance the proposed NUI
design. In addition to the aforementioned gestures, a tap gesture could activate regions of interest with additional
modalities of information such as text, images, and metadata. The user can interactively browse the different forms of
data by using intuitive touch gestures as shown in Fig.~\ref{fig:4}. This set of 2D interactions along with the 3D NUI design
presented earlier can compose an intuitive yet powerful workspace for an epigraphist who can now perform digitally
several parts of the epigraphic workflow.

@@

In this pilot study, a complete set of natural user interactions was designed based on the physical interactions of
epigraphists with real inscriptions. The proposed interactions utilize the existing sensors in a typical tablet
computer or smart phone in order to interactively relight a digitized inscription and manipulate the user's
perspective, using a set of intuitive gestures that imitate the natural interaction with a physical object. In the
proposed design, the epigraphist can ``hold'' a digital inscription, relight it by
reorienting it as a tangible object, observe it from different perspectives, and finally interact with other modalities
by following a set of 2D touch gestures. The prototype system was developed as part of the Digital Epigraphy and
Archaeology (DEA) project using the open-source library VisiNeat for 3D visualization and interaction and is compatible
with iOS, Android, and Microsoft RT tablet and smart phone devices. The interface is available through the web-site of
the project: http://www.digitalepigraphy.org

@@

We would like to acknowledge our collaborators for their invaluable interaction and feedback during the past three
years. We would also like to thank the UF College of the Arts and the UF Center for Greek Studies for providing
continuous funding support for this project.

>>>>

\noindent The technological advances in the last decade have equipped the general public with several handheld electronic
commodities that changed significantly daily routine in a personal and professional level and contributed to the user's
quality of life not only in developed countries but also in developing economies in Africa and Asia [Osman 2011].
Handheld devices, such as tablets and smart phones, not only connect the users with tremendous amount of information
through the internet, but also offer interfaces for natural user interaction that enable non-technology-oriented
populations to use computers intuitively.

@@

\noindent Understanding the users is one of the integral steps of interaction design, which is an iterative process during which
representative users interact with preliminary designs and provide useful feedback \citep{preece_interaction_2015}. For the
purposes of this study, our team interacted with early adopters of our prototype system, who were epigraphists and
conservation specialists from Cornell University, the University of California, Berkeley, the University of Lyon 2, the
Berlin-Bradenburg Academy of Sciences and Humanities, the U.K. National Archives, and the University of Florida. The
goals of our interaction were twofold: a) to study the various forms of physical interaction that epigraphists have
with an inscription as a real physical object and b) to expose epigraphists to a digital interface that imitates
their interaction routine, using digital replicas of physical objects.

@@

\noindent Natural User Interfaces consist of sensors that track the natural behavior of users and provide a natural form of
interactivity with computers and other electronic devices. The common forms of NUI sensors are: pressure sensors for
sensing touch gestures (e.g. touch screens and touch pads), motion sensors for sensing user-initiated changes in the
orientation and acceleration of the device (e.g. accelerometer, gyroscope, and compass), and position sensors for
tracking changes in the relative position of the user with respect to the device, such as body motions (e.g.
Microsoft's Kinect), finger motions (e.g. Occipital's Leap Motion), eye movements, and others.

An optimal interaction design solution should be intuitive, minimalistic, and non intrusive \citep{preece_interaction_2015}.
Therefore, in order to design interaction for epigraphy one should choose devices that are easily accessible by
epigraphists and do not interfere with their workspace (e.g. avoid introducing new devices or external sensors). All
forms of interaction described in Sec. 1.2 can be implemented, using motion and pressure sensors, which can be easily
found in tablet computers or smart phones. In both types of handheld devices the virtual object can also be assumed
handheld, without loss of generality, in order to generate a multi-sensory experience for the user (i.e. holding the
device = holding the digital inscription). Hence, NUI design is possible by utilizing accessible devices and
without the use of external sensors as it is described in details in the following sections.

@@

\noindent In order to achieve natural interactive relighting of an inscription, the system should imitate the process of
relighting a handheld physical object (such as a paper cast of inscription) by reorienting the object with respect to
the light of the environment. Without loss of generality, we can assume that the default virtual lighting source is
located on the ceiling, right above the device, which is also very intuitive choice as it is the most probable
real-world lighting condition. Under this assumption, a gyroscope, a sensor that tracks the orientation of the device
with respect to the gravitational vector, is enough to track the slope of the device with respect to the virtual light.
The top row of Fig.~\ref{fig:1} shows the approximated real-world orientation of a tablet computer as it was estimated using
the gyroscope of the device. The orientation is updated in real-time as the user moves the device.

@@

\noindent Another important form of interaction in the epigraphic routine is the change of the point of view in order to
understand better the structural details of the inscribed letterforms. Assuming that the model of the inscription is
parallel to the screen of the device, the change of point of view involves only change of the perspective projection of
the digital object without any virtual rotation. In such case the rotation of the object is equivalent to the physical
rotation of the device without any virtual rotation of the object.

@@

\noindent The forms of interactions presented in 1.3.1 and 1.3.2 involved only the motion sensors of a tabled computer without
utilizing the pressure sensors of the screen of the device. The commonly used touch gestures (such as 2-finger twist to
rotate, 2-finger pinch to zoom, and 2-finger translate to move) can be employed in order to enhance the proposed NUI
design. In addition to the aforementioned gestures, a tap gesture could activate regions of interest with additional
modalities of information such as text, images, and metadata. The user can interactively browse the different forms of
data by using intuitive touch gestures as shown in Fig.~\ref{fig:4}. This set of 2D interactions along with the 3D NUI design
presented earlier can compose an intuitive yet powerful workspace for an epigraphist who can now perform digitally
several parts of the epigraphic workflow.

@@

\noindent In this pilot study, a complete set of natural user interactions was designed based on the physical interactions of
epigraphists with real inscriptions. The proposed interactions utilize the existing sensors in a typical tablet
computer or smart phone in order to interactively relight a digitized inscription and manipulate the user's
perspective, using a set of intuitive gestures that imitate the natural interaction with a physical object. In the
proposed design, the epigraphist can ``hold'' a digital inscription, relight it by
reorienting it as a tangible object, observe it from different perspectives, and finally interact with other modalities
by following a set of 2D touch gestures. The prototype system was developed as part of the Digital Epigraphy and
Archaeology (DEA) project using the open-source library VisiNeat for 3D visualization and interaction and is compatible
with iOS, Android, and Microsoft RT tablet and smart phone devices. The interface is available through the web-site of
the project: http://www.digitalepigraphy.org

@@

\noindent We would like to acknowledge our collaborators for their invaluable interaction and feedback during the past three
years. We would also like to thank the UF College of the Arts and the UF Center for Greek Studies for providing
continuous funding support for this project.

### 2015-12-17 source/31/EAGLE2016_I Sicily_final copy.tex
no indent added

\section{Introduction: What is I.Sicily}

I.Sicily is an online, open access, digital corpus of the inscriptions on stone from ancient Sicily.\footnote{ The
corpus will be mounted at \url{www.csad.ox.ac.uk/sicily/isicily}, but is presently on a development server. A public face is
currently maintained via a blog at \url{http://isicily.wordpress.com/}, as well as on Facebook at \url{www.facebook.com/ISicily}
and on Twitter via \url{I.Sicily@Sicilyepigraphy}.} The corpus aims to include all texts inscribed on stone, in any language,
between approximately the seventh century BC and the seventh century AD. The corpus currently contains records for over
2,500 texts, and when complete is likely to contain c.4,000. The corpus is built upon a conversion from a legacy
dataset of metadata in MS Access to EpiDoc TEI XML.\footnote{\url{http://sourceforge.net/p/epidoc/wiki/Home/} [accessed
26.09.2015].} The XML records are held in an eXist database for xQuery access, and additionally indexed for full-text
search using SOLR/Lucene. The corpus and related information (museum list, bibliography) are published as Linked Data,
and are manipulated through a RESTful API. The records are queried and viewed through a web interface built with
AngularJS and jQuery javascript components. Mapping is provided in the browser by the Google Maps API, and ZPR (Zoom,
Pan, Rotate) image-viewing is provided by the IIIP image server and the OpenSeadragon javascript library.
%IIIP or IIIF ? I believe this is a typo

@@

\section{The motivations for and origins of I.Sicily}

The existing epigraphic landscape in Sicily is extremely diverse in two primary regards: on the one hand, the island has
a very mixed cultural and linguistic make-up, meaning that the epigraphic material is itself extremely varied, with
extensive use throughout antiquity of both Greek and Latin, as well as Oscan, Punic, Sikel, and Hebrew;\footnote{
Recent overview of much of the linguistic tradition in \citet{tribulato_language_2012}; and of the epigraphic material in \citet{gulletta_sicilia_1999}.} on the other hand, the publication of this material has a very uneven record and despite an excellent
pre-twentieth century tradition, the existing copora are far from complete and the ability of key journals such as SEG
or AE to keep pace with local publication has been limited.\footnote{ For an overview of the corpora tradition up to
the twentieth century, see \citet{de_vido_corpora_1999}.} A limited number of museum-based corpora have been published in recent
decades (for Catania, Palermo, Messina, and Termini Imerese, as well as the material from Lipari), but this has not
greatly improved the overall situation.

@@

\section{The aims of I.Sicily}

We outline briefly five areas in which I.Sicily aims to develop, facilitate and improve the study of epigraphic material
from ancient Sicily.

@@

Sicily is traditionally described as a `melting pot', the `crossroads of the Mediterranean'. The negative consequences
of the separation of epigraphic material according to linguistic traditions have recently been highlighted and directly
confronted by the Corpus Inscriptionum Iudaeae/Palaestinae (CIIP), edited by H. Cotton et al.\footnote{ Original notice
in \citet{cotton_corpus_1999}; presentation in \citet{cotton_corpus_2010}.} I.Sicily sets out to follow in that mould, since the
different linguistic traditions of Sicily not only exist side-by-side but interact constantly throughout the island’s
history, and no study of the epigraphic material can afford to ignore contemporary and parallel material in the other
languages.\footnote{ See e.g. \citet{manganaro_greco_1993}, \citet{prag_epigraphy_2002}, \citet{salmeri_i_2004}, \citet{korhonen_language_2011}, \citet{tribulato_language_2012}.} The situation
created by basic technologies such as Unicode and EpiDoc XML mean that there is now no reason not to be language
agnostic in the inclusion of material (the point may be obvious, but the tendency towards language-specific corpora is
still marked). The opportunities and possibilities offered by these technologies are considerable, even at the most
basic level, since, for example, searching can be made language specific or language neutral. One obvious area where
Sicilian studies are currently hampered by this partitioning is in the study of onomastics. The Lexicon of Greek
Personal Names records most instances of Greek names for the island, but Sicily is no less rich in non-Greek names
(Latin and others), and at present there is no onomasticon for the island.\footnote{ \url{www.lgpn.ox.ac.uk/} [accessed
26.09.2015].} Simply by the marking-up and indexing of all names in the island’s inscriptions, I.Sicily will generate a
powerful tool for future study. Although I.Sicily in its first phase is not undertaking morphological or syntactical
mark-up, the encoding of all these text in XML constitutes the necessary first stage in such a development, and we see
this as a highly desirable future project, and the possibilities for the field of historical linguistics are
considerable. The incorporation of a full range of metadata on the epigraphic support, geographical location,
chronology, etc. will likewise allow detailed analysis of cultural patterns and their relationship to language-use over
time.\footnote{ See Prag 2002 for a first effort in this direction.}

@@

Sicily presents a particularly extreme version of the common problem of identifying a text and its publication record.
No existing corpus in either Greek or Latin comes close to full coverage (CIL X and IG XIV are the largest individual
traditional corpora for the region, but both are over 125 years old and cover less than 30\% of the material now
known).\footnote{\citet{_inscriptiones_1883}; \citet{kaibel_inscriptiones_1890}.} Existing online databases improve on this situation, but the results
obtainable are of very varied value. The most comprehensive, in terms of the range of data recorded, is EDR (with which
I.Sicily is collaborating), which currently reports 1,906 records for `Sicilia'; but this reduces to 833 when limited
to texts on stone (`lapis' or `marmor'); contrast I.Sicily, with 2,563 records at the time of writing.\footnote{
\url{www.edr-edr.it/} [accessed 26.09.2015]} Clauss Slaby reports 4,374 records for `Sicilia' (including Christian
inscriptions, excluding `sigilla impressa'), but the return is inclusive of all kinds of epigraphic material, without
indication or discrimination, contains some duplication, is much harder to reconcile to existing records, and records
only text.\footnote{ \url{www.manfredclauss.de/} [accessed 26.09.2015].} The PHI database of Greek inscriptions has a rich
record of published Greek texts, but is text only and limited in outputs.\footnote{
\url{http://noapplet.epigraphy.packhum.org/regions/1156} [accessed 26.09.2015].} SEG references are available for 733
inscriptions on stone and AE references for 328 (data taken from the I.Sicily database and based upon comprehensive
manual trawls of SEG and AE).

@@

The traditional focus of epigraphic study upon the text, rather than the epigraphic support, means that epigraphic
publication in the past has frequently been relatively limited in the information which it has recorded about the
object on which the inscription is inscribed. This is a familiar complaint, and one which I.Sicily will address
wherever possible through full object description and a rich photographic record. However, a corollary of this general
problem is a very low level of information regarding current location and in particular the infrequent recording of
museum inventory numbers or similar information. This situation is inevitably exacerbated by the substantial (and very
positive) reorganisation and redevelopment of museum collections in Sicily recent decades – including a significant
increase in the number of museums and public collections.

@@

I.Sicily is actively generating rich geo-data for the individual inscriptions, both for the original findspot/provenance
and the current location (whether museum-based, on-site, or elsewhere), and we aim to provide map-based searching for
inscriptions, as well as text-based searching by ancient and modern place-names. In addition to full listing wherever
possible of both ancient and modern place names for epigraphic provenance, we are working to provide detailed location
information for each find-spot and the inscription’s current location, through a combination of library and map-based
research and the use of autopsy and GIS recording. At present geo-data is being recorded in two forms, both through the
use of explicit geographical locations in the form of longitude and latitude records in decimal degree form (using
{\textless}geo{\textgreater} elements), and through the use of Pleiades URI references wherever possible.\footnote{
\url{http://pleiades.stoa.org/} [accessed 26.09.2015].} We are committed to the long-term use of Pleiades as our primary
reference for ancient places, and to that end we aim to update and improve the Pleiades data for Sicilian locations, in
particular name data and sub-locations, in conjunction with the editing of the I.Sicily records.\footnote{ See e.g.
\citet{wilson_places:_2015}. Valeria Vitale (KCL) is currently undertaking a significant programme of data improvement in
Pleiades on behalf of I.Sicily; we are grateful to Tom Elliott and Jeffrey Becker for their support.}

@@

In order to support the aims outlined above, I.Sicily has taken a different approach to search and browse. Although
standard form-based search with paged results, like that of Google, makes sense for very large result sets, the
comparatively small number of records in I.Sicily (thousands versus the estimated 30 trillion web pages indexed in
Google) lends itself to a more direct and interactive approach - a spreadsheet/grid model (similar to Microsoft Excel)
that runs directly in the browser. Although it is tempting to repeat the standard web-form model, following the
argument that that’s what users expect, the spreadsheet approach will be much easier to use, narrowing quickly and
accurately to more easily interpreted results. Further, any subset of the spreadsheet, generated from interactive
filtering, can, with a single button push, be exported to CSV (comma separated values) for use outside I.Sicily. The
spreadsheet interacts particularly well with maps: all findspots or museums in a filtered subset of the grid can be
simultaneously shown on the map (see Fig.~\ref{fig:2}). The spreadsheet model also provides a very quick and intuitive (since so
many people are familiar with spreadsheets) means for editing records (in this case, inscriptions and museums) online.
This web-based spreadsheet model has only recently become feasible for the web, as web browsers have added more
functionality and new javascript libraries have been developed.

@@

As was extensively discussed at the first EAGLE conference (Paris 2014), the creation and availability of translations
is a major goal of the EAGLE project and its collaborators, and I.Sicily is no less committed to that
ambition.\footnote{ See Orlandi et al. 2014: Part II.} Translations are rarely available for any of the published
Sicilian inscriptions.\footnote{ French translations appear in Dubois 1989 and Dubois 2008.} It is obvious that the
inclusion of translations will make the material much more accessible to a wider audience both of students and the
general public. Equally, provision of translations will add to the value of the database as a resource for museums and
others curating the inscriptions recorded in the database. To that end, a long-term ambition of I.Sicily is to include
translations wherever possible in both English and Italian. We see this as one obvious area where public contribution
(`crowd-sourcing') will be invaluable (see below).

@@

The scale of the enterprise, and the available resources, mean that in its current form the project has limited itself
to inscriptions engraved on stone (the coverage of rupestral inscriptions/graffiti and of inscriptions painted on
stone/plaster is regrettably uneven). However, there is no reason in principle not to extend coverage in future to
include inscriptions on other materials. Similarly, although as noted above the current project does not include a
programme to mark up linguistic features of the texts, the commitment to the long-term maintenance of the corpus and
the open availability of the underlying XML records means that such a project would be entirely possible in the future.

@@

I.Sicily gratefully acknowledges the financial support of the John Fell Fund of the
University of Oxford, and of the Warden and Scholars of the House or College of Scholars of Merton in the University of
Oxford.

>>>>

\section{Introduction: What is I.Sicily}
\noindent I.Sicily is an online, open access, digital corpus of the inscriptions on stone from ancient Sicily.\footnote{ The
corpus will be mounted at \url{www.csad.ox.ac.uk/sicily/isicily}, but is presently on a development server. A public face is
currently maintained via a blog at \url{http://isicily.wordpress.com/}, as well as on Facebook at \url{www.facebook.com/ISicily}
and on Twitter via \url{I.Sicily@Sicilyepigraphy}.} The corpus aims to include all texts inscribed on stone, in any language,
between approximately the seventh century BC and the seventh century AD. The corpus currently contains records for over
2,500 texts, and when complete is likely to contain c.4,000. The corpus is built upon a conversion from a legacy
dataset of metadata in MS Access to EpiDoc TEI XML.\footnote{\url{http://sourceforge.net/p/epidoc/wiki/Home/} [accessed
26.09.2015].} The XML records are held in an eXist database for xQuery access, and additionally indexed for full-text
search using SOLR/Lucene. The corpus and related information (museum list, bibliography) are published as Linked Data,
and are manipulated through a RESTful API. The records are queried and viewed through a web interface built with
AngularJS and jQuery javascript components. Mapping is provided in the browser by the Google Maps API, and ZPR (Zoom,
Pan, Rotate) image-viewing is provided by the IIIP image server and the OpenSeadragon javascript library.
%IIIP or IIIF ? I believe this is a typo

@@

\section{The motivations for and origins of I.Sicily}
\noindent The existing epigraphic landscape in Sicily is extremely diverse in two primary regards: on the one hand, the island has
a very mixed cultural and linguistic make-up, meaning that the epigraphic material is itself extremely varied, with
extensive use throughout antiquity of both Greek and Latin, as well as Oscan, Punic, Sikel, and Hebrew;\footnote{
Recent overview of much of the linguistic tradition in \citet{tribulato_language_2012}; and of the epigraphic material in \citet{gulletta_sicilia_1999}.} on the other hand, the publication of this material has a very uneven record and despite an excellent
pre-twentieth century tradition, the existing copora are far from complete and the ability of key journals such as SEG
or AE to keep pace with local publication has been limited.\footnote{ For an overview of the corpora tradition up to
the twentieth century, see \citet{de_vido_corpora_1999}.} A limited number of museum-based corpora have been published in recent
decades (for Catania, Palermo, Messina, and Termini Imerese, as well as the material from Lipari), but this has not
greatly improved the overall situation.

@@

\section{The aims of I.Sicily}
\noindent We outline briefly five areas in which I.Sicily aims to develop, facilitate and improve the study of epigraphic material
from ancient Sicily.

@@

\noindent Sicily is traditionally described as a `melting pot', the `crossroads of the Mediterranean'. The negative consequences
of the separation of epigraphic material according to linguistic traditions have recently been highlighted and directly
confronted by the Corpus Inscriptionum Iudaeae/Palaestinae (CIIP), edited by H. Cotton et al.\footnote{ Original notice
in \citet{cotton_corpus_1999}; presentation in \citet{cotton_corpus_2010}.} I.Sicily sets out to follow in that mould, since the
different linguistic traditions of Sicily not only exist side-by-side but interact constantly throughout the island’s
history, and no study of the epigraphic material can afford to ignore contemporary and parallel material in the other
languages.\footnote{ See e.g. \citet{manganaro_greco_1993}, \citet{prag_epigraphy_2002}, \citet{salmeri_i_2004}, \citet{korhonen_language_2011}, \citet{tribulato_language_2012}.} The situation
created by basic technologies such as Unicode and EpiDoc XML mean that there is now no reason not to be language
agnostic in the inclusion of material (the point may be obvious, but the tendency towards language-specific corpora is
still marked). The opportunities and possibilities offered by these technologies are considerable, even at the most
basic level, since, for example, searching can be made language specific or language neutral. One obvious area where
Sicilian studies are currently hampered by this partitioning is in the study of onomastics. The Lexicon of Greek
Personal Names records most instances of Greek names for the island, but Sicily is no less rich in non-Greek names
(Latin and others), and at present there is no onomasticon for the island.\footnote{ \url{www.lgpn.ox.ac.uk/} [accessed
26.09.2015].} Simply by the marking-up and indexing of all names in the island’s inscriptions, I.Sicily will generate a
powerful tool for future study. Although I.Sicily in its first phase is not undertaking morphological or syntactical
mark-up, the encoding of all these text in XML constitutes the necessary first stage in such a development, and we see
this as a highly desirable future project, and the possibilities for the field of historical linguistics are
considerable. The incorporation of a full range of metadata on the epigraphic support, geographical location,
chronology, etc. will likewise allow detailed analysis of cultural patterns and their relationship to language-use over
time.\footnote{ See Prag 2002 for a first effort in this direction.}

@@

\noindent Sicily presents a particularly extreme version of the common problem of identifying a text and its publication record.
No existing corpus in either Greek or Latin comes close to full coverage (CIL X and IG XIV are the largest individual
traditional corpora for the region, but both are over 125 years old and cover less than 30\% of the material now
known).\footnote{\citet{_inscriptiones_1883}; \citet{kaibel_inscriptiones_1890}.} Existing online databases improve on this situation, but the results
obtainable are of very varied value. The most comprehensive, in terms of the range of data recorded, is EDR (with which
I.Sicily is collaborating), which currently reports 1,906 records for `Sicilia'; but this reduces to 833 when limited
to texts on stone (`lapis' or `marmor'); contrast I.Sicily, with 2,563 records at the time of writing.\footnote{
\url{www.edr-edr.it/} [accessed 26.09.2015]} Clauss Slaby reports 4,374 records for `Sicilia' (including Christian
inscriptions, excluding `sigilla impressa'), but the return is inclusive of all kinds of epigraphic material, without
indication or discrimination, contains some duplication, is much harder to reconcile to existing records, and records
only text.\footnote{ \url{www.manfredclauss.de/} [accessed 26.09.2015].} The PHI database of Greek inscriptions has a rich
record of published Greek texts, but is text only and limited in outputs.\footnote{
\url{http://noapplet.epigraphy.packhum.org/regions/1156} [accessed 26.09.2015].} SEG references are available for 733
inscriptions on stone and AE references for 328 (data taken from the I.Sicily database and based upon comprehensive
manual trawls of SEG and AE).

@@

\noindent The traditional focus of epigraphic study upon the text, rather than the epigraphic support, means that epigraphic
publication in the past has frequently been relatively limited in the information which it has recorded about the
object on which the inscription is inscribed. This is a familiar complaint, and one which I.Sicily will address
wherever possible through full object description and a rich photographic record. However, a corollary of this general
problem is a very low level of information regarding current location and in particular the infrequent recording of
museum inventory numbers or similar information. This situation is inevitably exacerbated by the substantial (and very
positive) reorganisation and redevelopment of museum collections in Sicily recent decades – including a significant
increase in the number of museums and public collections.

@@

\noindent I.Sicily is actively generating rich geo-data for the individual inscriptions, both for the original findspot/provenance
and the current location (whether museum-based, on-site, or elsewhere), and we aim to provide map-based searching for
inscriptions, as well as text-based searching by ancient and modern place-names. In addition to full listing wherever
possible of both ancient and modern place names for epigraphic provenance, we are working to provide detailed location
information for each find-spot and the inscription’s current location, through a combination of library and map-based
research and the use of autopsy and GIS recording. At present geo-data is being recorded in two forms, both through the
use of explicit geographical locations in the form of longitude and latitude records in decimal degree form (using
{\textless}geo{\textgreater} elements), and through the use of Pleiades URI references wherever possible.\footnote{
\url{http://pleiades.stoa.org/} [accessed 26.09.2015].} We are committed to the long-term use of Pleiades as our primary
reference for ancient places, and to that end we aim to update and improve the Pleiades data for Sicilian locations, in
particular name data and sub-locations, in conjunction with the editing of the I.Sicily records.\footnote{ See e.g.
\citet{wilson_places:_2015}. Valeria Vitale (KCL) is currently undertaking a significant programme of data improvement in
Pleiades on behalf of I.Sicily; we are grateful to Tom Elliott and Jeffrey Becker for their support.}

@@

\noindent In order to support the aims outlined above, I.Sicily has taken a different approach to search and browse. Although
standard form-based search with paged results, like that of Google, makes sense for very large result sets, the
comparatively small number of records in I.Sicily (thousands versus the estimated 30 trillion web pages indexed in
Google) lends itself to a more direct and interactive approach - a spreadsheet/grid model (similar to Microsoft Excel)
that runs directly in the browser. Although it is tempting to repeat the standard web-form model, following the
argument that that’s what users expect, the spreadsheet approach will be much easier to use, narrowing quickly and
accurately to more easily interpreted results. Further, any subset of the spreadsheet, generated from interactive
filtering, can, with a single button push, be exported to CSV (comma separated values) for use outside I.Sicily. The
spreadsheet interacts particularly well with maps: all findspots or museums in a filtered subset of the grid can be
simultaneously shown on the map (see Fig.~\ref{fig:2}). The spreadsheet model also provides a very quick and intuitive (since so
many people are familiar with spreadsheets) means for editing records (in this case, inscriptions and museums) online.
This web-based spreadsheet model has only recently become feasible for the web, as web browsers have added more
functionality and new javascript libraries have been developed.

@@

\noindent As was extensively discussed at the first EAGLE conference (Paris 2014), the creation and availability of translations
is a major goal of the EAGLE project and its collaborators, and I.Sicily is no less committed to that
ambition.\footnote{ See Orlandi et al. 2014: Part II.} Translations are rarely available for any of the published
Sicilian inscriptions.\footnote{ French translations appear in Dubois 1989 and Dubois 2008.} It is obvious that the
inclusion of translations will make the material much more accessible to a wider audience both of students and the
general public. Equally, provision of translations will add to the value of the database as a resource for museums and
others curating the inscriptions recorded in the database. To that end, a long-term ambition of I.Sicily is to include
translations wherever possible in both English and Italian. We see this as one obvious area where public contribution
(`crowd-sourcing') will be invaluable (see below).

@@

\noindent The scale of the enterprise, and the available resources, mean that in its current form the project has limited itself
to inscriptions engraved on stone (the coverage of rupestral inscriptions/graffiti and of inscriptions painted on
stone/plaster is regrettably uneven). However, there is no reason in principle not to extend coverage in future to
include inscriptions on other materials. Similarly, although as noted above the current project does not include a
programme to mark up linguistic features of the texts, the commitment to the long-term maintenance of the corpus and
the open availability of the underlying XML records means that such a project would be entirely possible in the future.

@@

\noindent I.Sicily gratefully acknowledges the financial support of the John Fell Fund of the
University of Oxford, and of the Warden and Scholars of the House or College of Scholars of Merton in the University of
Oxford.
